---
title: "Synergy master files merge"
author: "Calvin Ge"
date: "22/03/2021"
output: html_document
---

```{r setup} 

library(doBy)
library(sas7bdat)
library(data.table)
library(openxlsx)

base.path <- "/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/" 

output.path <-base.path

```


```{r lastest files}

load(file=paste0(base.path,"synergy_master_file_20210322.Rdata")) #master all.subjects file
load(file=paste0(base.path,"synergy_master_occ_hist_20210322.Rdata")) #master occ.hist file

#all.subjects2<-read.csv(file=paste0(base.path,"syn_all_subjects.txt"),sep="\t") #all.subjects text file 
#occ.hist2<-read.csv(file=paste0(base.path,"syn_all_occ_hist.txt"),sep="\t") #all.subjects text file 

```

```{r IARC meeting}

range(all.subjects$age[all.subjects$agegroup==1])
range(all.subjects$age[all.subjects$agegroup==2])
range(all.subjects$age[all.subjects$agegroup==3])
range(all.subjects$age[all.subjects$agegroup==4])
range(all.subjects$age[all.subjects$agegroup==5])
range(all.subjects$age[all.subjects$agegroup==6])
range(all.subjects$age[all.subjects$agegroup==7])
range(all.subjects$age[all.subjects$agegroup==8])
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==2])
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==3])
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==4]) #cut is same

table(all.subjects$smoking[],all.subjects$silica_qq0)



```



```{r trimmed file loading}

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/Silica_paper/Rdata/final.subjects.CG.Rdata") #clean file used in silica manuscript

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Asbestos/Rdata/final.subjects.linked.trimmed.Rdata")
load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Chromium/Rdata/final.subjects.linked.trimmed.Rdata")
load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Nickel/Rdata/final.subjects.linked.trimmed.Rdata")
load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/PAH/Rdata/final.subjects.linked.trimmed.Rdata")
load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Silica/Rdata/final.subjects.linked.trimmed.Rdata")

```

```{r columns to keep}

names(all.subjects)
colnames(all.subjects)[colnames(all.subjects) == 'duration'] <- 'smok_duration'
colnames(all.subjects)[colnames(all.subjects) == 'intensity'] <- 'smok_intensity'

keep<-all.subjects[,c("subjctid","study_name","country","sex","Enrolyr","Interyr","diagyr","Histotyp","status","age","agegroup",
                      "smoking","smok_intensity","smok_duration","packyears","packyrs","ever_smok","time_quit","time_since_quit",
                      "A","Blue","Years_employed","constr","miner","metalw","transp","veh_mech","farmer","since60","since70")]

colnames(keep)<-tolower(colnames(keep))
colnames(keep)[colnames(keep) == 'a'] <- 'list_a'


```

```{r merging all subjects files}

all.subjects.asb$subjctid<-sub("^0+([0-9]*)$", "\\1", all.subjects.asb$subjctid) #get rid of leading zero
all.subjects.asb$subjctid<-relevel(factor(all.subjects.asb$subjctid),ref="101031026")

all.subjects.cr6$subjctid<-sub("^0+([0-9]*)$", "\\1", all.subjects.cr6$subjctid) #get rid of leading zero
all.subjects.cr6$subjctid<-relevel(factor(all.subjects.cr6$subjctid),ref="101031026")

all.subjects.ni$subjctid<-sub("^0+([0-9]*)$", "\\1", all.subjects.ni$subjctid) #get rid of leading zero
all.subjects.ni$subjctid<-relevel(factor(all.subjects.ni$subjctid),ref="101031026")

all.subjects.pah$subjctid<-sub("^0+([0-9]*)$", "\\1", all.subjects.pah$subjctid) #get rid of leading zero
all.subjects.pah$subjctid<-relevel(factor(all.subjects.pah$subjctid),ref="101031026")

all.subjects.sili$subjctid<-sub("^0+([0-9]*)$", "\\1", all.subjects.sili$subjctid) #get rid of leading zero
all.subjects.sili$subjctid<-relevel(factor(all.subjects.sili$subjctid),ref="101031026")

keep$subjctid<-as.factor(as.character(keep$subjctid))

all.subjects<-merge(keep,all.subjects.asb,by = "subjctid", all.x = T)
all.subjects<-merge(all.subjects,all.subjects.cr6,by = "subjctid", all.x = T)
all.subjects<-merge(all.subjects,all.subjects.ni,by = "subjctid", all.x = T)
all.subjects<-merge(all.subjects,all.subjects.pah,by = "subjctid", all.x = T)
all.subjects<-merge(all.subjects,all.subjects.sili,by = "subjctid", all.x = T)

table(is.na(all.subjects$list_a)) #262
all.subjects$list_a[is.na(all.subjects$list_a)]<-0
all.subjects$blue[is.na(all.subjects$blue)]<-0
all.subjects$constr[is.na(all.subjects$constr)]<-0
all.subjects$miner[is.na(all.subjects$miner)]<-0
all.subjects$metalw[is.na(all.subjects$metalw)]<-0
all.subjects$transp[is.na(all.subjects$transp)]<-0
all.subjects$veh_mech[is.na(all.subjects$veh_mech)]<-0
all.subjects$farmer[is.na(all.subjects$farmer)]<-0

table(all.subjects$since60)
table(all.subjects$since70)

all.subjects$since60[is.na(all.subjects$since60)]<-0
all.subjects$since70[is.na(all.subjects$since70)]<-0
```

```{r all subject file checks}

### versus silica paper
table(all.subjects$ever_silica0,all.subjects$status) #matches paper
table(all.subjects$silica_qq0,all.subjects$status) #very close match to paper - slightly different qq cut
table(all.subjects$silica_dur_cat0, all.subjects$status)#matches paper
table(all.subjects$silica_time_since_exp40_0,all.subjects$status) #needs
table(all.subjects$smoking[all.subjects$status==0],all.subjects$ever_silica0[all.subjects$status==0])#matches paper
table(all.subjects$smoking[all.subjects$status==1],all.subjects$ever_silica0[all.subjects$status==1])#matches paper
range(all.subjects$silica_cum0[all.subjects$silica_qq0==1])
range(all.subjects$silica_cum0[all.subjects$silica_qq0==2])
range(all.subjects$silica_cum0[all.subjects$silica_qq0==3])
range(all.subjects$silica_cum0[all.subjects$silica_qq0==4])
table(all.subjects$histotyp,all.subjects$silica_qq0) #very close match with paper - slightly different qq cut


cumulative.0<- glm(status ~ 1 + factor(silica_qq0) + factor(study_name) + factor(agegroup) + sex + packyrs + factor(time_quit) + factor(list_a), data=all.subjects, family=binomial)
round(exp(cbind(summary(cumulative.0)$coefficients[2:5,1],confint(cumulative.0)[2:5,])),2) #very close match with paper

continuous.sili<-glm(status ~ 1 + silica_cum0 + factor(study_name) + factor(agegroup) + sex + packyrs + factor(time_quit) + factor(list_a), data=all.subjects, family=binomial)
summary(continuous.sili)$coefficients[2,] #5.827300e-02 8.282756e-03 7.035460e+00 1.986040e-12 exact match with paper model
round(exp(cbind(summary(continuous.sili)$coefficients[2:3,1],confint(continuous.sili)[2:3,])),3) #silica_cum0: 1.060 (1.04-1.077) per unit increase in cum. exposure
AIC(continuous.sili) #42035.84 exact match with paper model

### versus pahs draft
table(all.subjects$ever_pah0,all.subjects$status) #matches draft
table(all.subjects$pah_qq0,all.subjects$status) #very close match to draft
table(all.subjects$pah_dur_cat0, all.subjects$status)#matches draft
range(all.subjects$pah_cum0[all.subjects$pah_qq0==1])
range(all.subjects$pah_cum0[all.subjects$pah_qq0==2])
range(all.subjects$pah_cum0[all.subjects$pah_qq0==3])
range(all.subjects$pah_cum0[all.subjects$pah_qq0==4]) #same cut

cumulative.pah<- glm(status ~ 1 + factor(pah_qq0) + factor(study_name) + factor(agegroup) + sex + packyrs + factor(time_quit) + factor(list_a), data=all.subjects, family=binomial)
round(exp(cbind(summary(cumulative.pah)$coefficients[2:5,1],confint(cumulative.pah)[2:5,])),2) #very close match with draft

### versus asbestos paper
table(all.subjects$ever_asbestos0[all.subjects$sex==1],all.subjects$status[all.subjects$sex==1]) #a few dozen more unexposed subjects (both cases & controls) included here, however some subjects are missing from table 2 (9608+6802+3898+510 = 20818 controls and 6629+6958+2717+482 = 16786 cases in Table 2, actual # should be 20965/16901)
table(all.subjects$asbestos_qq0[all.subjects$sex==1],all.subjects$status[all.subjects$sex==1]) #very close in terms of exposed 
table(all.subjects$ever_asbestos0[all.subjects$sex==0],all.subjects$status[all.subjects$sex==0]) #also a few more controls here, which is likely the actual correct sum
table(all.subjects$asbestos_qq0[all.subjects$sex==0],all.subjects$status[all.subjects$sex==0]) #very close in terms of exposed
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==1])
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==2])
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==3])
range(all.subjects$asbestos_cum0[all.subjects$asbestos_qq0==4]) #cut is same

cumulative.asb<- glm(status ~ 1 + factor(asbestos_qq0) + factor(study_name) + factor(agegroup) + sex + packyrs + factor(time_quit) + factor(list_a), data=all.subjects, family=binomial)
round(exp(cbind(summary(cumulative.asb)$coefficients[2:5,1],confint(cumulative.asb)[2:5,])),2) #results very similar

table(all.subjects$status[all.subjects$sex==1],all.subjects$smoking[all.subjects$sex==1]) #as above. paper missing some subjects in all categories. diff not very large. 


```


```{r save all.subjects file}

save(all.subjects,file=paste0(output.path,"synergy_master_file_20210322.Rdata")) 

write.table(all.subjects, file=paste0(output.path,"syn_all_subjects.txt"), sep="\t", row.names=FALSE)

```


```{r load and merge occ.hist files}

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Asbestos/Rdata/unique.hist.asbestos.Rdata")
occ.hist.asb<-occ.hist[,c("subjctid","isco","jobno","asbestos","study_name","country","year","assignment.level","log.exposure","selected")]
colnames(occ.hist.asb)[colnames(occ.hist.asb) == 'log.exposure'] <- 'asbestos_log.exposure'

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Chromium/Rdata/unique.hist.chromium.Rdata")
occ.hist.cr6<-occ.hist[,c("subjctid","chromium","year","log.exposure")]
colnames(occ.hist.cr6)[colnames(occ.hist.cr6) == 'log.exposure'] <- 'chromium_log.exposure'

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Nickel/Rdata/unique.hist.nickel.Rdata")
occ.hist.ni<-occ.hist[,c("subjctid","nickel","year","log.exposure")]
colnames(occ.hist.ni)[colnames(occ.hist.ni) == 'log.exposure'] <- 'nickel_log.exposure'

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/PAH/Rdata/unique.hist.pah.Rdata")
occ.hist.pah<-occ.hist[,c("subjctid","pah","year","log.exposure")]
colnames(occ.hist.pah)[colnames(occ.hist.pah) == 'log.exposure'] <- 'pah_log.exposure'

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Silica/Rdata/unique.hist.silica.Rdata")
occ.hist.sili<-occ.hist[,c("subjctid","silica","year","log.exposure")]
colnames(occ.hist.sili)[colnames(occ.hist.sili) == 'log.exposure'] <- 'silica_log.exposure'

occ.hist<-merge(occ.hist.asb,occ.hist.cr6,by = c("subjctid","year"), all.x = TRUE)
occ.hist<-merge(occ.hist,occ.hist.ni,by = c("subjctid","year"), all.x = TRUE)
occ.hist<-merge(occ.hist,occ.hist.pah,by = c("subjctid","year"), all.x = TRUE)
occ.hist<-merge(occ.hist,occ.hist.sili,by = c("subjctid","year"), all.x = TRUE)

load(file="/Users/cbge/OneDrive - Universiteit Utrecht/SYNERGY/SYNJEM/Occ.hist/Rdata/all.subjects.CG.Rdata")  #need lagyr from this 
all.subjects<-all.subjects[all.subjects$excluded==0,] #37866 subjects remain
occ.hist <- merge(occ.hist,all.subjects[,c("subjctid","lagyr")],by="subjctid",all.x=TRUE) #1424594 lines
anyNA(occ.hist$lagyr)#hmmm
occ.hist <- subset(occ.hist,year<=lagyr) #this is to get all jobs before lagyr
nrow(occ.hist) #1396285

names(occ.hist)

#Note1: before use for calculating exposure by subjctid, exposure needs to be transformed form log scale back to normal
#Note2: unique ids in occ.hist is less than in all.subjects. the 262 subjects difference essentially have various d

save(occ.hist,file=paste0(output.path,"synergy_master_occ_hist_20210322.Rdata")) 
write.table(occ.hist, file=paste0(output.path,"syn_all_occ_hist.txt"), sep="\t", row.names=FALSE)
```

Important note 1: before use for calculating exposure by subjctid, exposure needs to be transformed form log scale back to normal. 

Important note 2: unique ids in occ.hist (n=37604) is less than in all.subjects (n=37866). The 262 subjects difference essentially have various neverworked/missing/blank/unknown jobs for their *entire* reported working lives. In our papers we have assumed that these 262 subjects have no work exposure and included them in the final epi analyses. 

Ongoing tasks: fill in the NAs in list.a with 0. All confirmed to be non-list a jobs. 
